---
title: "Catch-reports"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Catch-reports}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = F,
  message = F
)
```


To create catch reports first the data in the google sheet need to be up to date https://docs.google.com/spreadsheets/d/1uHmCuY3GXfSBCbsP61nAQeATBfoslXeS3PQNLlioWIk/edit#gid=0

Next open Rstudio, create a Project (name it something like `catch_reports`).  
I recommend then creating a folder by year `dir.create('2023')`.  
Install the `planteam` package and load it.

Overall the setup and running will look like:

```{r setup}
# install.packages("devtools")
# remotes::install_github("afsc-assessments/planteam")
library(planteam)
dir.create("2023")

catch_report(title = "1.B. Assessment of walleye pollock in the Bogoslof Island Region",
            authors = "James N. Ianelli, S. J. Barbeaux, Ivonne Ortiz, and D. McKelvey",
            stock = "pollock",
            area = "Bogoslof",
            year = 2023,
            next_full_year = 2024,
            catch_data_date = "November 5, 2023",
            output_dir = "2023")
            
```

This will generate a report that is placed in the selected output directory, so in this case it would be `2023/2023_Bogoslof_pollock_catch_report.docx`.

Similarly a catch report for BSAI other rockfish would be entered as:
```{r}
catch_report(title = "16. Assessment of the Other Rockfish stock complex in the Bering Sea/Aleutian Islands ",
             authors = "Jane Sullivan, Matt Callahan, Andy Kingham, Todd TenBrink, Ivonne Ortiz, Elizabeth Siddon, and Paul Spencer ",
             stock = "orox",
             area = "BSAI",
             year = 2023,
             last_full_year = 2022,
             next_full_year = 2024,
             catch_data_date = "November 5, 2023",
             output_dir = "2023")
```

If you have any problems please document them and post them as an issue on github at: https://github.com/afsc-assessments/planteam/issues

 
