---
params:
  doc_title:
  # authors:
  stock:
  area:
  year:
  last_full_year:
  next_full_year:
  link:  
  catch_data_date:
title: "`r params$doc_title`"

output:
  word_document:
     reference_docx: !expr system.file("template.docx", package = 'planteam')
     toc: false
     number_sections: false

author: "`r params$authors`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

flextable::set_flextable_defaults(font.family = "Times New Roman",
                                  font.size = 11)
```

######### Alaska Fisheries Science Center, National Marine Fisheries Service 

## Catch report

The most recent full assessment of this stock was in `r params$last_full_year` and can be found at the [Council website](https://www.npfmc.org/library/safe-reports/). 
Catch data are current through `r params$catch_data_date`. 
In accordance with the approved schedule, no assessment was conducted for this stock this year, however a full stock assessment will be conducted in `r params$next_full_year`. 
Until then, the values generated from the previous stock assessment will be rolled over for `r year+1` specifications.


```{r}
spp = params$stock
yr = year = params$year 
region = area = params$area
googlesheets4::gs4_auth()
  if(toupper(params$area)=="GOA"){
    df <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1uHmCuY3GXfSBCbsP61nAQeATBfoslXeS3PQNLlioWIk/edit?gid=0#gid=0",
                                    sheet = "goa")  %>% 
  dplyr::select(-c(completed, check, check_2))
  } else {
    df <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1uHmCuY3GXfSBCbsP61nAQeATBfoslXeS3PQNLlioWIk/edit?gid=0#gid=0",
                                  sheet = "bsai")  %>% 
  dplyr::select(-c(completed, checked))
  }

if(!(region %in% df$area) & region =="BSAI"){
  stop("the area must match the area field in the google sheet")
}

  if(area %in% c("GOA", "BSAI")){
    df %>% 
      dplyr::filter(species %in% spp,
                    year %in% (yr-1):(yr+2)) -> df 
    if(area=="GOA"){
      df %>% 
        dplyr::filter(area%in%c("Total", "GOA-wide")) -> df
    }
  } else {
    df %>% 
      dplyr::filter(species %in% spp,
                    year %in% (yr-1):(yr+2),
                    area %in% region) -> df 
  }

if(area=="GOA"){
  df %>% 
    dplyr::select(Year=year, OFL=ofl, ABC=abc, TAC=tac, Catch=catch) %>% 
    flextable::flextable() %>% 
    flextable::height(height=0.25) %>% 
    flextable::hrule(rule = "exact")  %>% 
      flextable::colformat_num(j=1, big.mark = "") %>% 
    # flextable::merge_v(j=1) %>% 
    # flextable::hline(i = rle(cumsum(.$body$spans$columns[,1]))$values) %>% 
    flextable::fix_border_issues() 
  } else {
  df %>% 
    dplyr::select(Area=area, Year=year, OFL=ofl, ABC=abc, TAC=tac, Catch=catch) %>% 
    flextable::flextable() %>% 
    flextable::height(height=0.25) %>% 
    flextable::hrule(rule = "exact")  %>%
        flextable::colformat_num(j=1, big.mark = "") %>% 
    flextable::merge_v(j=1) %>% 
    flextable::hline(i = rle(cumsum(.$body$spans$columns[,1]))$values) %>% 
    flextable::fix_border_issues() 
}

```
